NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-03-29 paper/background.tex
update background

\subsection{Natural Language Inference}
\label{sec:languageInference}
Natural Language Inference (NLI)~\cite{DaganRothSammons2013} is an important machine understanding task in NLP.
The goal of NLI is to predict the relationship between a premise (\textbf{P}) sentence and a hypothesis (\textbf{H}) sentence.
The prediction falls in one of three categories: \emph{entailment} (\textbf{E}), \emph{contradiction} (\textbf{C}), and \emph{neutral} (\textbf{N}).
A simple example is shown in Table~\ref{tab:NLI}.
In this case, the premise is ``Boy ate an apple''.
The hypothesis statement ``Kid ate fruits'' can be concluded from the premise. Therefore, the relationship between the premise and hypothesis is \emph{entailment}. However, we should note that such a relationship is not necessarily reversible. Since the concept ``fruit'' is less restrictive than that of ``apple'', we can not conclude ``Boy ate an apple'' from the statement ``Kid ate fruit''. The same logic applies to the hypothesis of ``Boy ate a Fuji apple.'', in which the premise neither imply nor oppose the hypothesis, therefore, their relationship is \emph{neutral}.
%\emph{Contradiction} means the textual of hypothesis opposes that of premise;
%and \emph{Neutral} implies no conclusion can be made on either \emph{Entailment} or \emph{Contradiction}.

@@

At the first glance, the task of natural language inference may seem less practical compared to other well-known NLP challenges such as machine translation; however, the ability to distinguish entailment and contradiction relationship is fundamental to understanding natural language at large.
%
Considering the ambiguousness of natural language and polysemy words, the inference task can become quite challenging (especially from the learning algorithm's point of view). Take the following sentences as an example (here \textbf{P} refers to premise, \textbf{H} refers to hypothesis):  (\textbf{P}) Facebook's IPO electrified the general public; (\textbf{H1}) Facebook went public; (\textbf{H2}) General Electric went public; (\textbf{H3}) People ignored Facebook's IPO. The literal similarity between ``Electric'' and ``electrified'' may trick a model to predict \textbf{H2} as \emph{entailment}. A model likely will also fail to understand the link between ``went public'' and ``IPO'', therefore, mistake \textbf{H1} as \emph{neutral}.
%
%The grand challenge in Natural Language Processing (NLP) is to have machine to acquire
%deep comprehension of textual information. Major tasks include Natural Language Inference,
%Machine Translation, Questions Answering, Summarization, and etc.
%Recently, neural networks models have performed strongly on these tasks.
%
Recently, Bowman et al. introduced a large corpus~\cite{BowmanAngeliPotts2015} for NLI tasks, which has helped spawn a new wave of effective neural network models for those tasks. Here, we focus on the analysis of the decomposable attention model~\cite{parikh2016emnlp} on this dataset.

@@

For natural language inference task, as illustrated in Fig.~\ref{fig:attention}, the attention information can be represented as a matrix describing the soft alignment between words in the premise and the words in the hypothesis. As we can see in this example, the higher values indicate significant alignments between words. The \emph{subject}, \emph{verb}, and \emph{object} between these two sentences are aligned correctly (words, such as ``an'', which has less importance in determine sentence relationship was assigned a lower score). Interestingly, the difference between the subject words does not affect their alignment. The classification stage can then utilize this information to determine the subject of the sentence are different, therefore, despite the other part of the sentences are identical, the relationship between the premise and the hypothesis is neutral.

>>>>

\subsection{Natural Language Inference}
\label{sec:languageInference}
Natural Language Inference (NLI)~\cite{DaganRothSammons2013} is an important machine understanding task in NLP.
The goal of NLI is to predict the relationship between a premise (\textbf{P}) sentence and a hypothesis (\textbf{H}) sentence.
The prediction falls in one of three categories: \emph{entailment} (\textbf{E}), \emph{contradiction} (\textbf{C}), and \emph{neutral} (\textbf{N}).
A simple example is shown in Table~\ref{tab:NLI}.
In this case, the premise is ``Boy ate an apple''.
The hypothesis statement ``Kid ate fruits'' can be concluded from the premise. Therefore, the relationship between the premise and hypothesis is \emph{entailment}. However, we should note that such a relationship is not necessarily reversible. Since the concept ``fruit'' is less restrictive than that of ``apple'', we cannot conclude ``Boy ate an apple'' from the statement ``Kid ate fruit''. The same logic applies to the hypothesis of ``Boy ate a Fuji apple.'', in which the premise neither implies nor opposes the hypothesis, and therefore, their relationship is \emph{neutral}.
%\emph{Contradiction} means the textual of hypothesis opposes that of premise;
%and \emph{Neutral} implies no conclusion can be made on either \emph{Entailment} or \emph{Contradiction}.

@@

At first glance, the task of natural language inference may seem less practical compared to other well-known NLP challenges such as machine translation; however, the ability to distinguish entailment and contradiction relationship is fundamental to understanding natural language at large.
%
Considering the ambiguousness of natural language and polysemy words, the inference task can become quite challenging (especially from the learning algorithm's point of view). Take the following sentences as an example (here \textbf{P} refers to premise, \textbf{H} refers to hypothesis):  (\textbf{P}) Facebook's IPO electrified the general public; (\textbf{H1}) Facebook went public; (\textbf{H2}) General Electric went public; (\textbf{H3}) People ignored Facebook's IPO. The literal similarity between ``Electric'' and ``electrified'' may trick a model to predict \textbf{H2} as \emph{entailment}. A model likely will also fail to understand the link between ``went public'' and ``IPO'', and therefore, mistake \textbf{H1} as \emph{neutral}.
%
%The grand challenge in Natural Language Processing (NLP) is to have machine to acquire
%deep comprehension of textual information. Major tasks include Natural Language Inference,
%Machine Translation, Questions Answering, Summarization, and etc.
%Recently, neural networks models have performed strongly on these tasks.
%
Recently, Bowman et al. introduced a large corpus~\cite{BowmanAngeliPotts2015} for NLI tasks, which has helped spawn a new wave of effective neural network models for those tasks. Here, we focus on the analysis of the decomposable attention model~\cite{parikh2016emnlp} on this dataset.

@@

For natural language inference task, as illustrated in Fig.~\ref{fig:attention}, the attention information can be represented as a matrix describing the soft alignment between words in the premise and the words in the hypothesis. As we can see in this example, the higher values indicate significant alignments between words. The \emph{subject}, \emph{verb}, and \emph{object} between these two sentences are aligned correctly (words, such as ``an'', which has less importance in determining sentence relationship, was assigned a lower score). Interestingly, the difference between the subject words does not affect their alignment. The classification stage can then utilize this information to determine the subject of the sentence is different, and therefore, despite the other part of the sentences are identical, the relationship between the premise and the hypothesis is neutral.
